\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-10 paper/main.tex
.

The way in which training and evaluation data is generated is related
to how the model is evaluated. An evaluation based on Macro AUC
evaluated the application scenario where a set of drugs and proteins
are given, and out of all possible pairs, the more likely interactions
need to be identified. However, this does not correspond to most
scenarios for drug repurposing where a drug that targets a {\em
  specific} protein (e.g., a protein involved in a disease) needs to
be identified. We use an evaluation measure based on micro-averaged
per protein (Micro AUC$_p$) to evaluate this application.

Finally, a potential source of differences in evaluation is how
negatives are treated during evaluation (and training).
Random sampling of unknown...

@@

We designed a
na\"ive predictor and find that it can achieve a performance only
slightly below state of the art drug--target prediction methods. We
investigated this finding further and

In the presence of imbalanced

Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it

Running split over proteins is harder than, drug and drug protein pair split (see below table)

this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)

Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug

>>>>

The way in which training and evaluation data is generated is related
to how the model is evaluated. An evaluation based on Macro AUC
evaluated the application scenario where a set of drugs and proteins
are given, and out of all possible pairs, the more likely interactions
need to be identified. However, this does not correspond to most
scenarios for drug repurposing where a drug that targets a {\em
  specific} protein (e.g., a protein involved in a disease) needs to
be identified. We use an evaluation measure based on micro-averages
per protein (Micro AUC$_p$) to evaluate this scenario, and we often
find substantial differences in predictive performance when evaluating
with Macro AUC and Micro AUC$_p$; generally, models that are trained
using a split over drug--target pairs perform worse in Micro AUC$_p$
than models that use a protein-based split, further providing evidence
that a drug--target split results in overfitting to dataset biases.

Finally, a potential source of differences in model performance is how
negatives are identified and treated during evaluation (and
training). There are few large sets of validated negative drug--target
interactions; consequently, many models (including \name{}) use all
unknown interactions as negatives. As there are many more negative
than positive interactions, negatives are then sub-sampled during
training resulting in a training set that is balanced between
positives and negatives (or a certain ratio is preserved). While this
is a reasonable strategy to deal with imbalanced data, it may be
problematic when the same sub-sampling is applied on the model's
evaluation set because it over-simplifies the evaluation process. The
performance differences is not usually visible when using ROC curves
but results in unrealistically high precision and therefore high area
under a precision-recall curve.

In summary, drug--target interaction prediction is not a single
computational problem in bioinformatics but a set of related
problems. Let $P$ be a set of proteins $P$ and $D$ a set of drugs; one
task can be to identify arbitrary pairs $(p,d)$ with $p \in P$ and
$d \in D$ that interact, another to identify a set of interacting
drugs for each $p \in P$, and yet another to identify a set of
proteins for each drug $d \in D$. The first task may be useful when no
particular drug or protein is considered; the second task when
searching for a drug that targets a specific (disease-associated)
protein; and the third when aiming to find new applications for a
given drug. The first task would best be evaluated using a Macro AUC,
the second and third using a Micro AUC$_p$ and Micro AUC$_d$.

@@

% We designed a
% na\"ive predictor and find that it can achieve a performance only
% slightly below state of the art drug--target prediction methods. We
% investigated this finding further and

% In the presence of imbalanced

% Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it

% Running split over proteins is harder than, drug and drug protein pair split (see below table)

% this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)

% Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

% microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
